---
title: "Learning Data Types with R"
author: "Binoy George"
date: "21 July 2018"
output: html_document
---

Knowing various Data Types in R:

Vectors
Lists
Matrices
Arrays
Factors
Data Frames


1) VECTOR DATA TYPE:
To create a VECTOR data type, we use c(). 'C' stands for 'combine' elements into a vector. Lets see some examples. 

```{r}

fruits <- c('apple','mango','grape','banana')
rand_vector <- c('john',9987,2+5,TRUE,'sam005')
num <- c(12,435,645645,678,-343,0.4,575)
bool <- c(1==1,3==3)
new_bool <- c(TRUE,TRUE,FALSE)

print(fruits)
print(rand_vector)
print(num)
print(bool)
print(new_bool)

print(class(fruits))
print(class(rand_vector))
print(class(num))
print(class(bool))
print(class(new_bool))

```


2) LIST DATA TYPE:
To create a LIST data type, we use list(). A list is an R-object which can contain many different types of elements inside it like vectors, functions and even another list inside it.. Lets see some examples. 

```{r}
l <- list(c(12,34,56,78,567),list(135,"exmpl",56.7),"char_ex",67,98,43)

print(l)

print(class(l))

```



3) MATRICES DATA TYPE:
A matrix is a two-dimensional rectangular data set. It can be created using a vector input to the matrix().. Lets see some examples. 

```{r}
m <- matrix(c(1:20),ncol = 10,nrow = 8)

n <- matrix(l,ncol = 10,nrow = 8)
#NOTE: 'l' is a list we created in the above R chunk.

print(m)
print(n)

class(m)
class(n)

```


4) ARRAYS DATA TYPE:
MATRICES are confined to two dimensions, ARRAYS can be of any number of dimensions. The array function takes a dim attribute which creates the required number of dimension.. Lets see some examples. 

```{r}

#NOTE: 'fruits' is a vector we created in the first R chunk.
arr1 <- array(fruits,dim =c(2,2,2))

arr2 <- array(fruits,dim =c(2,3,2))

arr3 <- array('NA',dim =c(2,2,3))

print(arr1)

print(arr2)

print(arr3)


print(class(arr1))
print(class(arr2))
print(class(arr3))

```


5) FACTOR DATA TYPE: 
FACTOR are created using VECTORS. FACTORS stores VECTORS along with their distinct values. FACTOR are created using the factor() function. The nlevels functions gives the count of levels or the distinct values of the elements that the FACTOR is storing. Lets see some examples.

```{r}
#Creating a VECTOR 'vv'
vv <- c(12,23,43,456,756,3423,321,12,12,123,4,3,4,4,4,4,2,3,5,6,7,3,22,6)

#Creating a FACTOR 'f'
f <- factor(vv)
print(f)
nlevels(f)

print(class(f))

```


6) 'DATA FRAMES' DATA TYPE: 
Data Frames are created using the data.frame() function. Data frames are tabular data objects. Unlike a matrix in data frame each column can contain different modes of data. The first column can be numeric while the second column can be character and third column can be logical. Lets see some examples.


```{r}

BMI <- data.frame(
  gender=c("Male","Female","Male","Male"),
  height=c(170,159,166,188),
  weight=c(80,88,70,45),
  age=c(32,39,40,44)
  )

print(BMI)

print(class(BMI))

```

